### 2022-09-06 latex/Chapters/DRLSearch.tex
update

@@

@@

\section{DL and DLR Heuristics}

So what is the link between all of these: A$^{*}$'s heuristics, \textbf{DL}, \textbf{RL}'s value iteration?
\\
\\
Sometimes we have ways of computing good solutions/heuristics to our puzzles, but cannot computationally do this for all possible states (maybe the state space is simply too large!). One approach in that case is to compute the solution/heuristic for \textit{some manageable number of} states, and let a \textbf{DL} model learn how to extrapolate to the rest of the state space. This is a case of \textit{supervised} learning in some sense, since we need a teacher-solver to tell us good solutions or heuristics to extrapolate from. In very simplified pseudo code, the procedure looks like the following:
\teal
\paragraph{}{\textbf{pseudo code -- \textbf{DL} heuristics}}
\begin{pseudocode}
###############################################################################
def deep_learning(puzzle_type,
                  teacher_heuristic,
                  loss_function,
                  network_architecture):
    """ pseudo code for learning a puzzle via deep learning and a teacher heuristic """
    target_value_function = {state: teacher_heuristic(state)
                             for state in generate_random_states(puzzle_type)}
    dl_heuristic = train_neural_network(target_value_function,
                                        loss_function,
                                        network_architecture)
    return dl_heuristic
###############################################################################
\end{pseudocode}
\black

@@

\section{Deep Q-Learning}

Another fundamental concept of \textbf{RL} is that of the Q-function $Q(s, a)$, the maximum expected reward the agent can obtain from state $s$, taking action $a$ and acting optimally afterwards. Notice that the knowledge of $V$ everywhere is equivalent to that of $Q$, since obviously:
\begin{equation} \label{eq:QI}
V(s) = max_{a \in \mathcal{A}(s)} Q(s, a)
\end{equation}
where $\mathcal{A}(s)$ are the transitions/actions available from state $s$. Conversely, we can recover $Q$ from $V$ since:
\begin{equation} \label{eq:QI2}
Q(s, a) = R(a) + V(a(s))
\end{equation}
where $R(a)$ is the reward of performing action $a$ and $a(s)$ is the state we reach by performing action $a$ from state $s$. Similarly to the way one can compute $V$ via value-iteration (at least in finite state and action spaces), $Q$ can also be computed by iteration (See \cite{Watkins1992} for details of convergence of that procedure).
\\
\\
Similarly to our discussion about \textbf{DRL}, in cases where the state space is large, it can be useful to approximate $Q$, or a somewhat related quantity (such as e.g. a probability distribution over the actions that can be taken from $s$) by a neural network, and use a similar procedure to that of the previous section to update the right-hand-side and left-hand-side of the iterations. That becomes \textbf{DQL}!
\\
\\
I have for this project followed the exact same procedure described in \cite{https://doi.org/10.48550/arxiv.1805.07470} (and also used by alpha-go \cite{AlphaGo}): the output of the neural network that is learnt by \textbf{DQL}, which actually is surprisingly similar to the approach and pseudo code shown for \textbf{DRL} earlier. The only 3 differences are that:
\begin{itemize}
\item the networks (target and current) now output a vector of size $a$ + 1, where $a$ is the number of possible actions/moves. That is, one dimension for each of the possible actions, representing the probability/desirability of that action, and one dimension for the value-function's value.
\item the target update by Q-iteration now needs to update both the value-function and the optimal decision ($a$-dimensional vector of 0s, with a 1 for the optimal decision only)
\item the loss function used need to be compatible with the size of the output and target vectors. I have used as loss the average of the mean square error of the value function difference (current network and target) and the cross entropy loss of the actions probability versus the actual optimal move.
\end{itemize}
Since in my \textbf{DRL} pseudo-code above (as well as in my actual code), the loss function is abstracted, and the network architecture (including the size of the output) configurable, the only modification to get \textbf{DQL} in lieu of \textbf{DRL} are extremely minimal, and essentially only to do with the Q-iteration-update section:

@@

Since my implementation of \textbf{DQL} is no longer just a cost-to-go estimate as with \textbf{DRL}, but a joint cost-to-go and distribution over actions (to be concrete, that means a 5-dimensional vector for the \textbf{SP} and a 13-dimensional vector for the \textbf{RC}). This leaves us with the question of how can we use the output of such a network to search for a solution. One simple answer (which I have tried, see e.g. subsection \ref{sec:S33DRLDQL}) is that we could disregard the distribution and only take into account the cost-to-go by simply using $A^{*}$. The rationale for doing so is that we might hope that jointly learning the cost-to-go and the actions via a combination of MSE and cross-entropy-loss produces a better network and leads to better heuristics, since we use more information during the learning and iterations.
\\
\\
Another possibility is to use an algorithm which combines both the actions distribution and the value function to inform its search. The Monte Carlo Tree Search (\textbf{MCTS}) is such an algorithm. In short, it generates at each step (and possibly in a distributed manner) paths to a \textit{leaf} on the frontier of unexpanded nodes, expands that node and checks if it is a goal state (and keeps going until the goal is found). The paths followed from initial state to leaves are decided via a combination of the probability distribution over actions (following the most promising actions) as well as the value function, according to a trade-off which can be tuned via a hyper-parameter. In order to prevent always following the same paths (which might take us very far from the goal), \textbf{MCTS} algorithms usually keep track of actions they have followed already and underweight their probability so that subsequent paths explore other actions. A pseudo code description of it looks like this:
\\
\teal
\paragraph{}{\textbf{pseudo code -- \textbf{MCTS}}}
\begin{pseudocode}
#################################################################
def MCTS(initial_node, time_out, q_v_network):
    """ pseudo code for Monte Carlo Tree Search """
    initialize_time_out(time_out)
    tree = Tree(initial_node)
    while True:
        check_time_out()
        leaf = construct_path_to_new_leaf(tree, q_v_network)
        if leaf.is_goal():
            return BFS(tree) # path to leaf can usually be improved by BFS
#################################################################
def construct_path_to_new_leaf(tree, q_v_network):
    leaf = tree.initial_node
    while leaf.expanded:
        actions_proba, value = q_v_network(leaf)
        historical_actions_count = leaf.actions_taken
        leaf = leaf.choose_child(actions_proba,
                                 value,
                                 historical_actions_count)
    # add children of leaf to tree
    leaf.expand()
    # penalize path taken to favour exploration
    leaf.increment_historical_actions_count()
    return leaf
#################################################################
\end{pseudocode}
\black

>>>>

\label{sec:TheoryBFSDFS}

@@

\label{sec:TheoryAStar}

@@

\section{DL and DLR Heuristics}
\label{sec:TheoryDLDRL}
So what is the link between all of these: A$^{*}$'s heuristics, \textbf{DL}, \textbf{RL}'s value iteration?
\\
\\
Sometimes we have ways of computing good solutions/heuristics to our puzzles, but cannot computationally do this for all possible states (maybe the state space is simply too large!). One approach in that case is to compute the solution/heuristic for \textit{some manageable number of} states, and let a \textbf{DL} model learn how to extrapolate to the rest of the state space. This is a case of \textit{supervised} learning in some sense, since we need a teacher-solver to tell us good solutions or heuristics to extrapolate from. In very simplified pseudo code, the procedure looks like the following:
\teal
\paragraph{}{\textbf{pseudo code -- \textbf{DL} heuristics}}
\begin{pseudocode}
###############################################################################
def deep_learning(puzzle_type,
                  teacher_heuristic,
                  loss_function,
                  network_architecture):
    """ pseudo code for learning a puzzle via deep learning and a teacher heuristic """
    target_value_function = {state: teacher_heuristic(state)
                             for state in generate_random_states(puzzle_type)}
    dl_heuristic = train_neural_network(target_value_function,
                                        loss_function,
                                        network_architecture)
    return dl_heuristic
###############################################################################
\end{pseudocode}
\black

@@

\section{Deep Q-Learning}
\label{sec:TheoryDQL}
Another fundamental concept of \textbf{RL} is that of the Q-function $Q(s, a)$, the maximum expected reward the agent can obtain from state $s$, taking action $a$ and acting optimally afterwards. Notice that the knowledge of $V$ everywhere is equivalent to that of $Q$, since obviously:
\begin{equation} \label{eq:QI}
V(s) = max_{a \in \mathcal{A}(s)} Q(s, a)
\end{equation}
where $\mathcal{A}(s)$ are the transitions/actions available from state $s$. Conversely, we can recover $Q$ from $V$ since:
\begin{equation} \label{eq:QI2}
Q(s, a) = R(a) + V(a(s))
\end{equation}
where $R(a)$ is the reward of performing action $a$ and $a(s)$ is the state we reach by performing action $a$ from state $s$. Similarly to the way one can compute $V$ via value-iteration (at least in finite state and action spaces), $Q$ can also be computed by iteration (See \cite{Watkins1992} for details of convergence of that procedure).
\\
\\
Similarly to our discussion about \textbf{DRL}, in cases where the state space is large, it can be useful to approximate $Q$, or a somewhat related quantity (such as e.g. a probability distribution over the actions that can be taken from $s$) by a neural network, and use a similar procedure to that of the previous section to update the right-hand-side and left-hand-side of the iterations. That becomes \textbf{DQL}!
\\
\\
I have for this project followed the exact same procedure described in \cite{https://doi.org/10.48550/arxiv.1805.07470} (and also used by alpha-go \cite{AlphaGo}): the output of the neural network that is learnt by \textbf{DQL}, which actually is surprisingly similar to the approach and pseudo code shown for \textbf{DRL} earlier. The only 3 differences are that:
\begin{itemize}
\item the networks (target and current) now output a vector of size $a$ + 1, where $a$ is the number of possible actions/moves. That is, one dimension for each of the possible actions, representing the probability/desirability of that action, and one dimension for the value-function's value.
\item the target update by Q-iteration now needs to update both the value-function and the optimal decision ($a$-dimensional vector of 0s, with a 1 for the optimal decision only)
\item the loss function used need to be compatible with the size of the output and target vectors. I have used as loss the average of the mean square error of the value function difference (current network and target) and the cross entropy loss of the actions probability versus the actual optimal move.
\end{itemize}
Since in my \textbf{DRL} pseudo-code above (as well as in my actual code), the loss function is abstracted, and the network architecture (including the size of the output) configurable, the only modification to get \textbf{DQL} in lieu of \textbf{DRL} are extremely minimal, and essentially only to do with the Q-iteration-update section:

@@

Since my implementation of \textbf{DQL} is no longer just a cost-to-go estimate as with \textbf{DRL}, but a joint cost-to-go and distribution over actions (to be concrete, that means a 5-dimensional vector for the \textbf{SP} and a 13-dimensional vector for the \textbf{RC}). This leaves us with the question of how can we use the output of such a network to search for a solution. One simple answer (which I have tried, see e.g. subsection \ref{sec:S33DRLDQL}) is that we could disregard the distribution and only take into account the cost-to-go by simply using $A^{*}$. The rationale for doing so is that we might hope that jointly learning the cost-to-go and the actions via a combination of MSE and cross-entropy-loss produces a better network and leads to better heuristics, since we use more information during the learning and iterations.
\\
\\
Another possibility is to use an algorithm which combines both the actions distribution and the value function to inform its search. The Monte Carlo Tree Search (\textbf{MCTS}) is such an algorithm. In short, it generates at each step (and possibly in a distributed manner) paths to a \textit{leaf} on the frontier of unexpanded nodes, expands that node and checks if it is a goal state (and keeps going until the goal is found). The paths followed from initial state to leaves are decided via a combination of the probability distribution over actions (following the most promising actions) as well as the value function, according to a trade-off which can be tuned via a hyper-parameter. In order to prevent always following the same paths (which might take us very far from the goal), \textbf{MCTS} algorithms usually keep track of actions they have followed already and underweight their probability so that subsequent paths explore other actions. A pseudo code description of it looks like this:
\\
\teal
\paragraph{}{\textbf{pseudo code -- \textbf{MCTS}}}
\begin{pseudocode}
#################################################################
def MCTS(initial_node, time_out, q_v_network):
    """ pseudo code for Monte Carlo Tree Search """
    initialize_time_out(time_out)
    tree = Tree(initial_node)
    while True:
        check_time_out()
        leaf = construct_path_to_new_leaf(tree, q_v_network)
        if leaf.is_goal():
            return BFS(tree) # path to leaf can usually be improved by BFS
#################################################################
def construct_path_to_new_leaf(tree, q_v_network):
    leaf = tree.initial_node
    while leaf.expanded:
        actions_probas, actions_values = q_v_network(leaf)
        """ choose next node based on joint actions probas & values
        as well as history of actions taken
        """
        leaf = leaf.choose_child(actions_probas,
                                 actions_values)
    # add children of leaf to tree
    leaf.expand()
    # penalize path taken to favour exploration
    leaf.increment_historical_actions_count()
    return leaf
#################################################################
\end{pseudocode}
\black

### 2022-09-06 latex/Chapters/Implementation.tex
update

\chapter{Code} % Main chapter title

\label{Chapter3} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

@@

\section{rubiks.heuristics}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebaseheuristics}
%\decoRule
\caption[Codebase]{rubiks.heuristics}
\label{fig:Codebaseheuristics}
\end{figure}
\label{HSS}
This module contains base class Heuristic, also a Factory. Heuristic can instantiate the following heuristics, which we can use in the AStar strategy from the previous section:
\begin{itemize}
\item \textbf{Manhattan}: This heuristic is specific to the SP. It simply adds for each tile (other than the empty tile) the $L_{1}$ distance between their current position and their target position. We can quickly see that this heuristic is admissible. Indeeed, each move in the SP moves one (and one only) tile by one position. In other puzzles, it is generally the case that a move will affect the position of many tiles (e.g. RC). Therefore, if tiles could somehow freely move on top of one-another (that is, we remove the constraint that there can be at most one tile per compartment, the number of moves necessary to solve the SP would exactly be the Manhattan distance.
\\
\\
The reason why it is interesting to have a known admissible heuristic is that we can obvioulsy compare other heuristics (DL, DRL, DQL, etc) in terms of optimality.
\\
\\
I have also implemented an improvement to the Manhattan distance, which I shall call Manhattan++ in here (and in code logs and graphs) and can be activated by simply passing \textit{plus=True} to the Manhattan Heuristic (see example \ref{ASSS} as well as a thorougher performance comparison on the (n=2, m=5) SP in \ref{MHComp}). It is based on the concept of linear constraints that I read about in lecture notes from Washington University (\cite{SlidingPuzzleLectureNotes}). The idea is simply that when tiles are already placed on their target row or column, but not in the expected order, we will need to get the tiles out of the way of one another to get to the goal state, and this is not accounted for by the Manhattan distance. For instance, in the following (n=2, m=3) SP, the Manhattan distance does not account for the fact that, at best, tile \red 3 \black needs to get out of the way for tiles \blue 1 \black and \blue 2 \black to move, and then needs to get back in its row, adding a cost of 2 to the Manhattan distance.

@@

Two important things to notice are that linear constraints across rows and columns (which I might generically refer to as \textit{lines} in the following) can to be added without breaking admissibility (hence giving more pessimistic, or accurate, cost estimates than simple Manhattan). This is because if a tile is involved in two linear constraints, it will need to get out of the way both horizontally and vertically. The second thing is that when several pairs in a line are not in order, we cannot simply add 2 for each distinct out-of-order pair, the right penalty to add is more subtle than that and needs to be computed recursively. For instance, let us now consider the following configuration:
\begin{center}
\begin{five}
\setrow{2}{\red 3, \red 2, \red 1}
\setrow{1}{4, 5, 0}
\end{five}
\end{center}
The correct penalty to add is not 6 (3 times 2 since all pairs (1, 2), (1, 3) and (2, 3) are out of order, but only 4. Indeed if, say, tile 3 got somehow out of the way at the back of the SP (imagine just another dimension there where we can move tiles) and tile 2 got out of the way by moving down to let tile 1 pass across, we could be done by simply adding 4 to the Manhattan distance (2 to move tile 3 out and back, and 2 to move tile 2 out and back). The correct way to compute the penalty cost for linear constraints is therefore to do it recursively, taking the minimum additional cost of moving either left-most or right-most tile of the line under consideration out of the way (that additional cost to move these left-most or right-most tile is 2 if not at their expected order in the line, 0 otherwise) plus the penalty of reordering the rest of the line.
\\
\\
Finally, as suggested by the reference lecture notes (which give very vague details about the above subtleties), I have precomputed all the penalties for all possible rows, columns and all possible tiles ordering they could have and saved the corresponding penalties in a database. For memory efficiency, I also only saved penalties which are non-zero. The very first time any call to Manhattan++ is made, for a given dimension (n, m), the appropriate data base is computed and populated.
\\
\\
Notice that for an (n, m) SP, there are n rows, each of which can have $\frac{(n * m)!}{(n * m - m)!}$ different ordering of tiles and m columns which can have n columns, each can have $\frac{(n * m)!}{(n * m - n)!}$ different ordering of tiles. This means the pre-computations and data-base sizes for the Manhattan++ heuristic are actually manageable, as it grows much slower than the number of possible puzzles. The maximum number of penalties to compute for $(n, m) \leq (5, 5)$ are:

@@

\item \textbf{PerfectHeuristic}: this reads from a data base the optimal costs, pre-computed by the PerfectLearner (see below \ref{PLcode})
\item \textbf{DeepLearningHeuristic}: this uses a network which has been trained using DRL by the DeepReinforcementLearner (see below \ref{DRLcode})
\end{itemize}

\section{rubiks.deeplearning}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebasedeeplearning}
%\decoRule
\caption[Codebase]{rubiks.deeplearning}
\label{fig:Codebasedeeplearning}
\end{figure}
This module is a wrapper around Pytorch. It contains:
\begin{itemize}
\item \textbf{DeepLearning}: a Puzzled Loggable Factory that can instantiate some configurable deep networks, and provide the necessary glue with the rest of the code base so that puzzles be seemlessly passed to the networks and trained on. 
\item \textbf{FullyConnected}: wrapper around a Pytorch fully connected network, with configurable hidden layers and size. There are some params as well to add drop out, and to indicate whether or not the inputs are one hot encoding (in which case the first layer is automatically adjusted in size, using information from the puzzle dimension).
\item \textbf{Convolutional}: similar wrapper to FullyConnected, but with the ability to add some parallel convolutional layers to complement fully connected layers.
\end{itemize}

\section{rubiks.learners}
\label{sec:codelearners}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebaselearners}
%\decoRule
\caption[Codebase]{rubiks.learners}
\label{fig:Codebaselearners}
\end{figure}
\label{PLcode}
\label{DRLcode}
This module implements learners, which learn something from a puzzle, store what they learnt, and can display interesting things about what they learnt.

\begin{itemize}
\item \textbf{Learner} is a Puzzled Loggable Factory. It provides some common code to learners (to save or purge what they learnt), kick off learning and plot results. Concrete derived implementation define what and how they learn, and what interesting they can display about this learning process. Currently the two implemented learners are:
\item \textbf{PerfectLearner}: It instantiates an optimal solver ($A^{*}$ with a configurable heuristic - but will only accept heuristic that advertise themselves as optimal. The learning consists in generating all the possible configuration of the considered puzzle, solve them with the optimal solver, and save the optimal cost of it as well as those of the whole solution path. The code allows for parallelization, stop and restart so that we can run on several different occasions and keep completing a database of solutions if necessary or desired. Once the PerfectLearner has completed its job, it can display some interesting information, such as the puzzle's God's number, the distribution of number of puzzles versus optimal cost, the hardest configuration it came across, and how long it took it to come up with the full knowledge of that puzzle. I will show in section \ref{PLSS} how to run an example. Notice that for puzzles of too high dimension, where my computing resources will not allow to solve exhaustively all the configurations of a given dimension, this class can still be used to populate a data base of optimal costs, which can then be used by DeepLearner. If it is to be used this way, the PerfectLearner can be configured to use perfectly random configurations to learn from, rather than going through the configurations one by one in a well defined order.

\item \textbf{DeepLearner} tbd
\item \textbf{DeepReinforcementLearner}: It instantiates a DeepLearning (network), and trains it using DRL. It then saves the trained network, which can then be used in the DeepLearningHeuristic we have seen earlier in section \ref{HSS}. The DeepReinforcementLearner runs for a number of epochs (or less if, based on other parameters discussed below, it is deemed to have converged).

@@

\section{rubiks.solvers}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebasesolvers}
%\decoRule
\caption[Codebase]{rubiks.solvers}
\label{fig:Codebasesolvers}
\end{figure}
This module implements solvers, which solve puzzles. The base class Solver is a Factory of solvers, and in addition to being able to instantiating the following types of solvers, can run different solvers through a similal sequences of random puzzles (for various increasing degrees of difficulty (shuffling), and/or perfectly shuffled ones) and display a comparison of how they perform in a number of metrics.

\begin{itemize}
\item \textbf{DFSSolver} TBD
\item \textbf{BFSSolver} TBD
\item \textbf{AStarSolver} TBD
\item \textbf{NaiveSlidingSolver} TBD
\item \textbf{Kociemba} TBD
\item \textbf{MonteCarloSearchTreeSolver} TBI I want to implement this in August when working on the Rubiks' ... following the Rubik's paper fromMcAleer \& Agostilenni et al \cite{https://doi.org/10.48550/arxiv.1805.07470}
\end{itemize}

>>>>

\chapter{Implementation} % Main chapter title

\label{sec:Implementation} % Change X to a consecutive number; for referencing this chapter elsewhere, use \ref{ChapterX}

@@

\section{rubiks.heuristics}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebaseheuristics}
%\decoRule
\caption[Codebase]{rubiks.heuristics}
\label{fig:Codebaseheuristics}
\end{figure}
\label{HSS}
This module contains base class Heuristic, also a Factory. Heuristic can instantiate the following heuristics, which we can use in the AStar strategy from the previous section:
\begin{itemize}
\item \textbf{Manhattan}: This heuristic is specific to the SP. It simply adds for each tile (other than the empty tile) the $L_{1}$ distance between their current position and their target position. We can quickly see that this heuristic is admissible. Indeeed, each move displaces one (and one only) tile by one position. In other puzzles, it is generally the case that a move will affect the position of many tiles (e.g. RC). Therefore, if tiles could somehow freely move on top of one-another (that is, we remove the constraint that there can be at most one tile per compartment, the number of moves necessary to solve the \textbf{SP} would exactly be the Manhattan distance.
\\
\\
The reason why it is interesting to have a known admissible heuristic is that we can obvioulsy compare other heuristics (DL, DRL, DQL, etc) in terms of optimality.
\\
\\
I have also implemented an improvement to the Manhattan distance, which I shall call Manhattan++ in here (and in code logs and graphs) and can be activated by simply passing \textit{plus=True} to the Manhattan Heuristic (see example \ref{ASSS} as well as a thorougher performance comparison on the 2x5 \textbf{SP} in \ref{MHComp}). It is based on the concept of linear constraints that I read about in lecture notes from Washington University (\cite{SlidingPuzzleLectureNotes}). The idea is simply that when tiles are already placed on their target row or column, but not in the expected order, we will need to get the tiles out of the way of one another to get to the goal state, and this is not accounted for by the Manhattan distance. For instance, in the following 2x3 puzzle, the Manhattan distance does not account for the fact that, at best, tile \red 3 \black needs to get out of the way for tiles \blue 1 \black and \blue 2 \black to move, and then needs to get back in its row, adding a cost of 2 to the Manhattan distance.

@@

Two important things to notice are that linear constraints across rows and columns (which I might generically refer to as \textit{lines} in the following) can be added without breaking admissibility (hence giving more pessimistic, or accurate, cost estimates than simple Manhattan). This is because if a tile is involved in two linear constraints, it will need to get out of the way both horizontally and vertically. The second thing is that when several pairs in a line are not in order, we cannot simply add 2 for each distinct out-of-order pair, the right penalty to add is more subtle than that and needs to be computed recursively. For instance, let us now consider the following configuration:
\begin{center}
\begin{five}
\setrow{2}{\red 3, \red 2, \red 1}
\setrow{1}{4, 5, 0}
\end{five}
\end{center}
The correct penalty to add is not 6 (3 times 2 since all pairs (1, 2), (1, 3) and (2, 3) are out of order, but only 4. Indeed if, say, tile 3 got somehow out of the way at the back of the SP (imagine just another dimension there where we can move tiles) and tile 2 got out of the way by moving down to let tile 1 pass across, we could be done by simply adding 4 to the Manhattan distance (2 to move tile 3 out and back, and 2 to move tile 2 out and back). The correct way to compute the penalty cost for linear constraints is therefore to do it recursively, taking the minimum additional cost of moving either the left-most or the right-most tile of the line under consideration out of the way (that additional cost to move these left-most or right-most tile is 2 if not at their expected order in the line, 0 otherwise) plus the penalty of reordering the rest of the line.
\\
\\
Finally, as suggested by the reference lecture notes (which give very vague details about the above subtleties), I have precomputed all the penalties for all possible rows, columns and all possible tiles ordering they could have and saved the corresponding penalties in a database. For memory efficiency, I also only saved penalties which are non-zero. The very first time any call to Manhattan++ is made, for a given dimension nxm, the appropriate linear constraint penalties are computed and populated in a database.
\\
\\
Notice that for an nxm \textbf{SP}, there are n rows, each of which can have $\frac{(n * m)!}{(n * m - m)!}$ different ordering of tiles and m columns which can each have $\frac{(n * m)!}{(n * m - n)!}$ different ordering of tiles. This means the pre-computations and data-base sizes for the Manhattan++ heuristic are actually manageable, as it grows much slower than the number of possible puzzles. The maximum number of penalties to compute for $nxm \leq 5x5$ are:

@@

\item \textbf{PerfectHeuristic}: this reads from a data base the optimal costs, pre-computed by the PerfectLearner (see below \ref{PLcode})
\item \textbf{DeepLearningHeuristic}: this uses a network which has been trained using \textbf{DL} (DeepLearner) \textbf{DRL} (DeepReinforcementLearner) or \textbf{DQL} (DeepQLearner). See \ref{DRLcode} for a discussion of these three learners.
\item \textbf{DeepQLearningHeuristic}: this uses a network which has been trained using \textbf{DQL} by the DeepQLearner (see below \ref{DQLcode}). The DeepQLearningHeuristic follows the same interface as the other heuristics, and therefore can be used by A$^{*}$ for instance. It also has an additional method \textit{optimal\_actions} that returns the learnt probability distribution over actions for a given puzzle, for use in search algorithms (e.g. \textbf{MCTS}) which make use of this to inform their search.
\end{itemize}

\section{rubiks.deeplearning}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebasedeeplearning}
%\decoRule
\caption[Codebase]{rubiks.deeplearning}
\label{fig:Codebasedeeplearning}
\end{figure}
This module is a wrapper around Pytorch. It contains:
\begin{itemize}
\item \textbf{DeepLearning}: a Puzzled Loggable Factory that can instantiate some configurable deep networks, and provide the necessary glue with the rest of the code base so that puzzles be seemlessly passed to the networks and trained on. 
\item \textbf{FullyConnected}: wrapper around a Pytorch fully connected network, with configurable hidden layers and size. There are some params as well to add drop out, and to indicate whether or not the inputs are one hot encoding (in which case the first layer is automatically adjusted in size, using information from the puzzle dimension). There is also importantly a \textit{joint\_policy} parameter to indicate whether we want the output to be 1 dimensional (e.g. for value function learning) or $a+1$-dimensional, where $a$ is the size of the actions space (for joint policy-value learning).
\item \textbf{Convolutional}: similar wrapper to FullyConnected, but with the ability to add some parallel convolutional layers to complement fully connected layers.
\end{itemize}

\section{rubiks.learners}
\label{sec:codelearners}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebaselearners}
%\decoRule
\caption[Codebase]{rubiks.learners}
\label{fig:Codebaselearners}
\end{figure}
\label{PLcode}
\label{DRLcode}
\label{DQLcode}
This module implements learners, which learn something from a puzzle, store what they learnt, and can display interesting things about what they learnt.

\begin{itemize}
\item \textbf{Learner} is a Puzzled Loggable Factory. It provides some common code to learners (to save or purge what they learnt), kick off learning and plot results. Concrete derived implementation define what and how they learn, and what interesting they can display about this learning process. The four implemented learners are:
\item \textbf{PerfectLearner}: It instantiates an optimal solver ($A^{*}$ with a configurable heuristic - but will only accept heuristic that advertise themselves as optimal. The learning consists in generating all the possible configuration of the considered puzzle, solve them with the optimal solver, and save the optimal cost of it as well as those of the whole solution path. The code allows for parallelization, stop and restart so that we can run on several different occasions and keep completing a database of solutions if necessary or desired. Once the PerfectLearner has completed its job, it can display some interesting information, such as the puzzle's God's number, the distribution of number of puzzles versus optimal cost, the hardest configuration it came across, and how long it took it to come up with the full knowledge of that puzzle. I will show in section \ref{PLSS} how to run an example. Notice that for puzzles of too high dimension, where my computing resources will not allow to solve exhaustively all the configurations of a given dimension, this class can still be used to populate a data base of optimal costs, which can then be used by DeepLearner. If it is to be used this way, the PerfectLearner can be configured to use perfectly random configurations to learn from, rather than going through the configurations one by one in a well defined order.

\item \textbf{DeepLearner} DeepLearner instanciates a DeepLarning (neural network), and trains it on training data by minimizing the mean square error between the cost-to-go from the target solver versus the neural network. The main parameters driving this learner are:
\begin{itemize}
\item \textit{training\_data\_from\_data\_base} which controls whether the training samples are generated on the fly via another solver (defaults to A$^{*}[Manhattan++]$ for \textbf{SP} and A$^{*}[Kociemba]$ for \textbf{RC}) or read from a data base (similarly constructed from other solvers, but has the advantage of being done offline)
\item \textit{nb\_epochs} and \textit{threshold} which control exit criteria for the network training.
\item \textit{nb\_sequences}, \textit{nb\_shuffles\_min}, \textit{nb\_shuffles\_max}, \textit{training\_data\_freq} and \textit{training\_data\_every\_epoch} which control how often we regenerate new training data, how many sequences of puzzles it contais, how scrambled the puzzles should be.
\item \textit{learning\_rate}, \textit{scheduler}, \textit{optimizer} and assorted parameters which control the optimisation at each backward propagation. 
\end{itemize}
Once it has completed (or on a regular basis), DeepLearner saves the trained network, which can then be used with DeepLearningHeuristic to guide A$^{*}$.
\item \textbf{DeepReinforcementLearner}: It instantiates a DeepLearning (network), and trains it using \textbf{DRL} essentially following the pseudo-code from \ref{sec:TheoryDLDRL}. Unlike DeepLearner, the targets are here generated via value iteration rather than via another solver (teacher). It has very similar parameters to the DeepLearner, in particular some parameters to control exit criteria (same as DeepLearner's plus a few that help it exit early when the value function's range has not been increasing anymore for a while), some parameters to control and configure the optimiser, some to control the training puzzles (how many, how often, how scrambled).

\item \textbf{DeepQLearner}: The implementation of the DeepQLearner is actually only a few lines of code, for it inherits pretty much all of its behaviour from the DeepReinforcementLearner. The only overwritten functions are:

\begin{itemize}
\item \textit{get\_loss\_function} which constructs the loss function to be used by the network's training. Instead of returning a simple torch.nn.MSELoss, it returns a function which computes the average of torch.nn.MSELoss on the value-function and torch.nn.CrossEntropyLoss on the actions.
\item \textit{\_\_construct\_target\_\_} which performs the the Q-V-iteration to update the targets, as described in pseudo code form in \ref{sec:TheoryDQL}. This function updates not only the value function of a node given its children nodes, but also updates the optimal action vector (0 for non optimal actions, 1 for the optimal action leading to the child node of highest value function).
\end{itemize}

@@

\section{rubiks.solvers}
\label{sec:codesolvers}
\begin{figure}[H]
\centering
\includegraphics[scale=0.25]{./Figures/codebasesolvers}
%\decoRule
\caption[Codebase]{rubiks.solvers}
\label{fig:Codebasesolvers}
\end{figure}
This module implements solvers, which solve puzzles. The base class Solver is a Factory of solvers, and in addition to being able to instantiating the following types of solvers, can run different solvers through a similal sequences of random puzzles (for various increasing degrees of difficulty (shuffling), and/or perfectly shuffled ones) and display a comparison of how they perform in a number of metrics (chosable from percentage of puzzles solved before time out, percentage of optimal puzzle (if optimality can be ascertained), mean-median-maximum run time, mean-median-maximum expanded nodes (for solvers where it makes sense), mean-median-max cost, and an optimality score which compares the total sum of cost over all puzzles versus a benchmark solver). Ideally the benchmark solver to compute the optimality score should be one which is provably optimal, but in the case of the \textbf{RC} that was not possible so I used A$^{*}$ with KociembaHeuristic as my benchmark. For the \textbf{SP} I used the optimal A$^{*}$ with Manhattan++ heuristic.

\begin{itemize}
\item \textbf{DFSSolver} This solver is literally a few lines of code, extremely close to the pseudo code detailed in \ref{sec:TheoryBFSDFS}
\item \textbf{BFSSolver} Ditto, see \ref{sec:TheoryBFSDFS}
\item \textbf{AStarSolver} Ditto, see \ref{sec:TheoryAStar}
\item \textbf{NaiveSlidingSolver} The NaiveSolver is conceptually simple, but was far from trivial to implement and debug. Its functioning is detailed in the appendix section \ref{sec:Examples}. Basically, it does what a beginner player would do (certainly what I do when I play the sliding puzzle), solving the top row first, then the left column, hence reducing the puzzle to a smaller dimensional one, and keep iterating until done.
\item \textbf{KociembaSolver} This solver is a wrapper around two libraries. The first one, hkociemba (see \cite{HKociemba}) is a 2x2x2 implementation written by Kociemba himself, the second one is a Python pip-installable which solves the 3x3x3 Rubik's. My  wrapper simply translates my cube representation in that accepted by these 2 solvers, and their result string into moves usable by my code. I also do some additional massaging for the 3x3x3 case, due to a shortcoming/bug in the 3x3x3 implementation, as I have detailed in appendix \ref{sec:Kociemba}.
\item \textbf{MonteCarloSearchTreeSolver} TBI I want to implement this in August when working on the Rubiks' ... following the Rubik's paper fromMcAleer \& Agostilenni et al \cite{https://doi.org/10.48550/arxiv.1805.07470}
\end{itemize}
